---
title: "Introduction to STL Forecasting"
output: html_notebook
---

In this notebook we present a [decomposition model](https://fabletools.tidyverts.org/reference/decomposition_model.html) that combines STL (Seasonal and Trend decomposition using Loess) and ETS/ARIMA with [tidyverts](https://tidyverts.org/). From the documentation:

> This function allows you to specify a decomposition combination model using any additive decomposition. It works by first decomposing the data using the decomposition method provided to dcmp_fn with the given formula. Secondary models are used to fit each of the components from the resulting decomposition.

For more details see Forecasting: [Principles and Practice, Section 3.6 STL Decomposition](https://otexts.com/fpp3/stl.html).

```{r}
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(fable)
library(feasts)
library(stringr)
options(dplyr.summarise.inform=F) 
```

# Read Data

We will use the Basel temperature data set.

```{r}
raw_df <- read_csv('../../data/basel_weather.csv')
```

# EDA

```{r}
data_df <- raw_df %>%
  rename(temperature    = `Basel Temperature [2 m elevation corrected]`,
         precipitation  = `Basel Precipitation Total`,
         wind_speed     = `Basel Wind Speed [10 m]`,
         wind_direction = `Basel Wind Direction [10 m]`) %>%
  mutate(date      = date(timestamp),
         year      = year(timestamp),
         month     = month(timestamp),
         day       = day(timestamp),
         dayofyear = yday(timestamp),
         hour      = hour(timestamp))

daily_data_df <- data_df %>%
  group_by(date, year, month, day, dayofyear) %>%
  summarise(temperature = mean(temperature)) %>%
  as_tsibble(index=date)

daily_data_df %>% head()
```

```{r}
autoplot(daily_data_df, temperature) +
  labs(title='Basel Temperature (Daily)', y=expression(degree*C))
```
The time series contains a strong seasonal component.

```{r}
daily_data_df %>% gg_season(temperature)
```
We check the decomposition of the time series using STL.

```{r}
daily_data_df %>% 
  model(STL(temperature ~ season(period = 365, window = Inf))) %>% 
  components() %>% 
  autoplot()
```

# Train-Test Split

```{r}
train_test_cut_date <- as_date('2019-01-01')

df_train <- daily_data_df %>% filter(date < train_test_cut_date)
df_test <- daily_data_df %>% filter(date >= train_test_cut_date)

daily_data_df %>%
  mutate(data_set = if_else(date < train_test_cut_date, 'train', 'test')) %>%
  ggplot(aes(x=date, y=temperature, color=data_set)) +
  geom_line() +
  labs(title='Basel Temperature (Daily)', y=expression(degree*C)) +
  geom_vline(xintercept = train_test_cut_date, linetype = "longdash")
```
# Model Fit

We fit an exponential smoothing and an ARIMA model to the seasonal adjusted time series.

```{r}
fit <- df_train %>%
  model(
    stl_arima = decomposition_model(
      STL(temperature ~ season(period = 365, window = Inf)),
      ARIMA(season_adjust ~ 0 + pdq(2, 1, 1) + PDQ(0, 0, 0))
    ),
    stl_ets =  decomposition_model(
      STL(temperature ~ season(period = 365, window = Inf)),
      ETS(season_adjust ~ season("N"))
    )
  )
```

# Generate Forecast

```{r}
fc <- fit %>%
  forecast(h = nrow(df_test))
```

```{r}
error <- fc %>% accuracy(df_test)

rmse_arima <- error %>% filter(.model=="stl_arima") %>% pull(RMSE)
rmse_ets <- error %>% filter(.model=="stl_ets") %>% pull(RMSE)
```

```{r}
fc %>% autoplot(df_test, level=c()) +
  labs(title='Basel Temperature (Daily)', y=expression(degree*C)) +
  scale_color_discrete(labels = c(stl_arima = str_interp("STL+ARIMA rmse = $[.2f]{rmse_arima}"),
                                  stl_ets = str_interp("STL+ETS rmse = $[.2f]{rmse_ets}")))
```

