---
title: "Introduction To Exponential Smoothing"
output: html_notebook
---

Loading packages and set up the notebook.

```{r}
library(readr)
library(dplyr)
library(tibble)
library(ggplot2)
library(fable)
```

Load sample data (prepared earlier).

```{r}
df <- read_csv("../../data/intro_ets.csv") %>%
  rowid_to_column("id") %>%
  as_tsibble(index=id)
df %>% head()
```

# Simple Exponential Smoothing

Formula for simple exponential smoothing...

\begin{align}
s_t &= \alpha x_t + (1-\alpha) s_{t-1} \\
    &= \alpha x_t + (1-\alpha)(\alpha x_{t-1} + (1-\alpha) s_{t-2}) \\
    &= \alpha x_t + (1-\alpha) \alpha x_{t-1} + (1-\alpha)^2 s_{t-2}
\end{align}

Lets now try out simple exponential smoothing on a time series

```{r}
df %>% autoplot(t1)
```
Train the model with arbitrary smoothing_level (alpha) parameter:

```{r}
fit <- df %>%
  model(
    ETS(t1 ~ trend("N", alpha=0.2) + season("N"))
  )
```

The plot below shows the original time series and the backfit exponential smoothing values. Adjusting the alpha parameter controls the smoothness. Values of alpha near 1 put larger weights on more recent values. Values of alpha near 0 put larger weights on older values and hence we get a smoother fit.

```{r}
fit %>%
  augment() %>%
  pivot_longer(c(t1, .fitted)) %>%
  ggplot(aes(x=id, y=value, colour=name)) +
  geom_line()
```
Now lets look at the forecast:

```{r}
fit %>%
  forecast(h=20) %>%
  autoplot(df)
```
What happens when we try it on a a series with a trend?

```{r}
df %>% autoplot(t2)
```
```{r}
df %>%
  model(
    ETS(t2 ~ trend("N", alpha=0.2) + season("N"))
  ) %>%
  forecast(h=20) %>%
  autoplot(df)
```
# Double Exponential Smoothing

\begin{align}
s_t &= \alpha x_t + (1-\alpha)(s_{t-1} + b_{t-1}) \\
b_t &= \beta (s_t-s_{t-1}) + (1-\beta) b_{t-1}
\end{align}

```{r}
df %>%
  model(
    ETS(t2 ~ trend("A", alpha=0.5, beta=0.5) + season("N"))
  ) %>%
  forecast(h=20) %>%
  autoplot(df)
```
Ok, that works on simple series, but what if we want to exploit a known seasonal pattern?

```{r}
df %>% autoplot(t3)
```
# Triple Exponential Smoothing

Add in a another formula to account for periodic cycles!

Triple Exponential Smoothing (Additive)

\begin{align}
s_0 &= x_0 \\
s_t &= \alpha(x_t - c_{t-L})+(1-\alpha)(s_{t-1}+b_{t-1}) \\
b_t &= \beta(s_t - s_{t-1}) + (1-\beta)b_{t-1} \\
c_t &= \gamma(x_t - s_{t-1} - b_{t-1}) + (1 - \gamma)c_{t-L} \\
F_{t+m} &= s_t + mb_t + c_{t-L+1+(m-1) \bmod L} \\
\end{align}

```{r}
df %>%
  model(
    ETS(t3 ~ trend("A", alpha=0.4, beta=0.01) + season("A", period=20, gamma=0.2))
  ) %>%
  forecast(h=20) %>%
  autoplot(df)
```
# Exponential Growth

```{r}
df %>% autoplot(t4)
```
```{r}
df %>%
  model(
    ETS(t4 ~ trend("M", alpha=0.6, beta=0.5) + season("N"))
  ) %>%
  forecast(h=20, times=100) %>%
  autoplot(df)
```
# Choosing alpha, beta, gamma automatically

The fable package can find the best parameters of alpha, beta and gamma via backfitting. Once you have chosen the appropriate model above, then this a great option to get a baseline.

```{r}
fit <- df %>%
  model(
    ETS(t3 ~ trend("A") + season("A", period=20))
  )
fit %>% coef()
```

```{r}
fit %>%
  forecast(h=20) %>%
  autoplot(df)
```
 
 
 
