---
title: "Causal Arima Intro"
editor_options:
  markdown:
    wrap: 88
---

## Excercise

On October 4, 2018 the Florence branch of an Italian supermarket chain introduced a
price policy change on several store-brand products by permanently reducing their price.
Your goal is to estimate the causal effect of the price policy change on the sales of
one of the addressed products.

-   Data

The data set includes the daily sales of one product from September 2017 to April 2019,
its price and information on national holidays and already built-in day-of-the-week
dummies.

Dates : vector of dates Sales : time series of daily sales counts (number of units sold)
TruePrice : the true unit price of the product Price : the new unit price series, set to
a constant value in the post-intervention period hol : holiday dummy taking value 1
before and after a national Italian holiday dec.sun : dummy taking value 1 on December
Sundays sat--thr : day-of-the-week dummies

-   Questions

    1.  What is the total number of units sold due to the price reduction after 1-week,
        1 month and 3 months from the intervention?

    2.  What is the total number of units sold at the end of the analysis period?

    3.  Plot the causal effect and residual diagnostics

    4.  Assume a multiplicative effect, how much more did the product sold due to the
        permanent price reduction?

-   R Package

The suggested R package is `CausalArima`, you can install the development version from
GitHub with

```{r}
# install.packages("devtools")
# devtools::install_github("FMenchetti/CausalArima")
```

## Prepare Notebook

```{r}
library("CausalArima")
library("tidyverse")
```

## Read Data

```{r}
raw_data_df <- read_csv(file = "../../data/btsa-causalarima_2021-11-02_2033/dataset.csv")

raw_data_df %>% head()
```

## EDA

```{r}
data_df <- raw_data_df %>% 
  mutate(
    price_change =  (TruePrice != Price),
    days_since_price_change = cumsum(as.numeric(price_change))
  )
```

```{r}
# date on which prices changed
price_change_date <- data_df %>%
  filter(price_change) %>%
  pull(Dates) %>%
  min()
```

```{r}
ggplot(data=data_df) +
  geom_line(mapping = aes(x = Dates, y = Sales), color="black") +
  geom_vline(xintercept = price_change_date, color = "red", linetype="dashed") +
  ggtitle(label = "Daily Sales")
```

```{r}
ggplot(data=data_df) +
  geom_line(mapping = aes(x = Dates, y = TruePrice,  color="TruePrice")) +
  geom_line(mapping = aes(x = Dates, y = Price, color="Price")) +
  geom_vline(xintercept = price_change_date, color = "black", linetype="dashed") +
  scale_color_brewer(palette="Set1") +
  labs(title="Price over time", color="price") 
```

## Answers

    1.  What is the total number of units sold due to the price
        reduction after 1-week, 1 month and 3 months from the
        intervention?

```{r}
compute_sales_after_price_change <- function ( .df , day_window ) {
  
  .df_filtered <- .df %>% filter(price_change & days_since_price_change <= day_window)
  
  sales_day_window <- .df_filtered %>% 
    summarise(Sales = sum(Sales, na.rm = TRUE)) %>% 
    pull(Sales)
  
  return( sales_day_window )
}

compute_sales_after_price_change(.df = data_df, day_window = 7)
```

```{r}
day_windows <- c(7, 30, 90, 120, 180)

tibble(
  day_windows = day_windows,
  sales = day_windows %>% map_dbl(function (x) compute_sales_after_price_change(.df = data_df, day_window = x))
) %>% 
ggplot() + 
  geom_line(mapping = aes(x = day_windows, y = sales), color="blue") +
  geom_point(mapping = aes(x = day_windows, y = sales), color="blue") +
  labs(title="Sales since price change")

```

    2.  What is the total number of units sold at the end of the
        analysis period?

```{r}
total_number_sold_end <- compute_sales_after_price_change(
  .df = data_df,
  day_window = (data_df %>% filter(price_change) %>% nrow())
)

total_number_sold_end
```

    3.  Plot the causal effect and residual diagnostics

```{r}
ggplot(data=data_df) +
  geom_line(mapping = aes(x = Dates, y = log(Sales)), color="black") +
  geom_vline(xintercept = price_change_date, color = "red", linetype="dashed") +
  ggtitle(label = "Daily Sales - Log Scale")
```


```{r}

y <- ts(data =(data_df %>% pull(Sales)), frequency = 1)
log_y <- log(y)
dates <- data_df %>% pull(Dates)
x_reg <- data_df %>% select(hol, dec.sun, sat, sun, mon, tue, wed, thr)

ce <- CausalArima(
  y = log_y, 
  auto = TRUE,
  dates = dates,
  int.date = price_change_date, 
  xreg = x_reg 
)
```

```{r}
forecasted <- plot(ce, type = "forecast")

forecasted
```

```{r}
impact_p <- plot(ce, type  ="impact")

grid.arrange(impact_p$plot, impact_p$cumulative_plot)
```
```{r}
ce$model$coef
```

    4.  Assume a multiplicative effect, how much more did the product sold due to the
        permanent price reduction?
        
Let us denote by $y_{i}$ and $\hat{y}_{i}$ the observed and the forecasted time series of the arima model. Hence, the *causal effect* is given by:

$$
ce_{i} = \log(y_{i}) - \log(\hat{y}_{i}) = \log\left(\frac{y_{i}}{\hat{y}_{i}}\right) 
$$
Therefore 

$$
\exp(ce_{i}) = \frac{y_{i}}{\hat{y}_{i}}  \Longrightarrow \hat{y}_{i} = \frac{y_{i}}{\exp(ce_{i})}
$$


```{r}
multiplicative_causal_effect <- ce$causal.effect  %>% exp()

price_changed_data_df <- data_df %>% 
  filter(price_change) %>% 
  mutate(
    multiplicative_causal_effect = multiplicative_causal_effect,
    sales_estimated_no_price_change = Sales / multiplicative_causal_effect,
    causal_effect = Sales - sales_estimated_no_price_change
  )

ggplot(data = price_changed_data_df) +
  geom_line(mapping = aes(x = Dates, y = multiplicative_causal_effect), color = "purple") +
  labs(title = "Multiplicative Causal Effect From C-ARIMA Model")
```
```{r}
causal_effect <- price_changed_data_df %>%
  summarise(causal_effect = sum(causal_effect, na.rm = TRUE)) %>% 
  pull(causal_effect)
```

```{r}
print(glue::glue(
"In the change of price period there were {total_number_sold_end} from where we estimate {round(causal_effect)} were caused by the price change. This accounts for {round( 100 * causal_effect / total_number_sold_end, digits=2)}% of the sales during this period."
))
```
