---
title: "Causal Arima Intro"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Excercise

On October 4, 2018 the Florence branch of an Italian supermarket chain
introduced a price policy change on several store-brand products by
permanently reducing their price. Your goal is to estimate the causal
effect of the price policy change on the sales of one of the addressed
products.

-   Data

The data set includes the daily sales of one product from September 2017
to April 2019, its price and information on national holidays and
already built-in day-of-the-week dummies.

Dates : vector of dates Sales : time series of daily sales counts
(number of units sold) TruePrice : the true unit price of the product
Price : the new unit price series, set to a constant value in the
post-intervention period hol : holiday dummy taking value 1 before and
after a national Italian holiday dec.sun : dummy taking value 1 on
December Sundays sat--thr : day-of-the-week dummies

-   Questions

    1.  What is the total number of units sold due to the price
        reduction after 1-week, 1 month and 3 months from the
        intervention?

    2.  What is the total number of units sold at the end of the
        analysis period?

    3.  Plot the causal effect and residual diagnostics

    4.  Assume a multiplicative effect, how much more did the product
        sold due to the permanent price reduction?

-   R Package

The suggested R package is `CausalArima`, you can install the
development version from GitHub with

```{r}
install.packages("devtools")
devtools::install_github("FMenchetti/CausalArima")
```

## Prepare Notebook

```{r}
library("CausalArima")
library("tidyverse")
```

## Read Data

```{r}
raw_data_df <- read_csv(file = "../../data/btsa-causalarima_2021-11-02_2033/dataset.csv")

raw_data_df %>% head()
```

## EDA

```{r}
data_df <- raw_data_df %>% 
  mutate(
    price_change =  (TruePrice != Price),
    days_since_price_change = cumsum(as.numeric(price_change))
  )
```

```{r}
# date on which prices changed
price_change_date <- data_df %>%
  filter(price_change) %>%
  pull(Dates) %>%
  min()
```

```{r}
ggplot(data=data_df) +
  geom_line(mapping = aes(x = Dates, y = Sales), color="black") +
  geom_vline(xintercept = price_change_date, color = "red", linetype="dashed") +
  ggtitle(label = "Daily Sales")
```

```{r}
ggplot(data=data_df) +
  geom_line(mapping = aes(x = Dates, y = TruePrice,  color="TruePrice")) +
  geom_line(mapping = aes(x = Dates, y = Price, color="Price")) +
  geom_vline(xintercept = price_change_date, color = "black", linetype="dashed") +
  scale_color_brewer(palette="Set1") +
  labs(title="Price over time", color="price") 
```

## Answers

    1.  What is the total number of units sold due to the price
        reduction after 1-week, 1 month and 3 months from the
        intervention?

```{r}
compute_sales_after_price_change <- function ( .df , day_window ) {
  
  return( .df %>% 
    filter(price_change & days_since_price_change <= day_window) %>% 
    summarise(Sales = sum(Sales, na.rm = TRUE)) %>% 
    pull(Sales)
  )
}

compute_sales_after_price_change(.df = data_df, day_window = 7)
```

```{r}
day_windows <- c(7, 30, 90)

tibble(
  day_windows = day_windows,
  sales = day_windows %>% map_dbl(function (x) compute_sales_after_price_change(.df = data_df, day_window = x))
) %>% 
ggplot() + 
  geom_line(mapping = aes(x = day_windows, y = sales), color="blue") +
  geom_point(mapping = aes(x = day_windows, y = sales), color="blue") +
  labs(title="Sales since price change")

```

    2.  What is the total number of units sold at the end of the
        analysis period?

```{r}
compute_sales_after_price_change(.df = data_df, day_window = (data_df %>% filter(price_change) %>% nrow()))
```

    3.  Plot the causal effect and residual diagnostics

```{r}
ce <- CausalArima(
  y = log(ts(data_df %>% pull(Sales))), 
  auto = TRUE,
  dates = (data_df %>% pull(Dates)),
  int.date = price_change_date, 
  xreg = (data_df %>% select(hol, dec.sun, sat, sun, mon, tue, wed, thr)), 
)
```


```{r}
forecasted <- plot(ce, type="forecast")
forecasted
```
```{r}
impact_p <- plot(ce, type  ="impact")
grid.arrange(impact_p$plot, impact_p$cumulative_plot)
```

